NanoMaLy

### 2015-11-16 paper/pldi16-submission/interactive.tex
add syntax for substitutions

\subsection{Interactive Semantics}
\label{sec:inter-semant}
%
\begin{figure*}[t]
\relDescription{Trace Syntax}
$$
\begin{array}{rrcl}
  & \tr & ::= & \bullet \spmid \singlestep{e}{e}; \tr \spmid \subterm{e}{e}; \tr
\end{array}
$$
\\
\relDescription{\subtermssym}
\begin{gather*}
\begin{array}{lcl}
\subterms{\eapp{e_1}{e_2}}   & \defeq & \subterm{\eapp{e_1}{e_2}}{e_1}; \subterm{\eapp{e_1}{e_2}}{e_2} \\
\subterms{\eplus{e_1}{e_2}}   & \defeq & \subterm{\eplus{e_1}{e_2}}{e_1}; \subterm{\eplus{e_1}{e_2}}{e_2} \\
\subterms{\eif{e_1}{e_2}{e_3}}   & \defeq & \subterm{\eif{e_1}{e_2}{e_3}}{e_1}; \\
                                &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_2}; \\
                                &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_3} \\
\subterms{\elet{x}{e_1}{e_2}}   & \defeq & \subterm{\elet{x}{e_1}{e_2}}{e_1}; \\
                                &        & \subterm{\elet{x}{e_1}{e_2}}{e_2} \\
\subterms{\efun{x}{e}}       & \defeq & \subterm{\efun{x}{e}}{e} \\
\subterms{e}                 & \defeq & \bullet
\end{array}
\end{gather*}
\judgementHead{Evaluation}{\stepg{e}{\su}{\tr}{e}{\su}{\tr}}
\begin{gather*}
\inference[\recontext]
  {\stepg{e}{\su}{\tr}{e_1}{\su_1}{\tr_1}}
  {\stepg{C[e]}{\su}{\tr}{C[e_1]}{\su_1}{\singlestep{C[e]}{C[e_1]}; \subterms{C[e_1]}; \tr_1}}
\\ \\
\inference[\reappgood]
  {\pair{\efun{x}{e}}{\su_2} = \force{v_1}{\tfun{\thole{}}{\thole{}}}}
  {\stepg{\eapp{v_1}{v_2}}{\su_1}{\tr}
         {e\sub{x}{v_2}}{\su_1\su_2}{\singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}; \subterms{e\sub{x}{v_2}}; \tr}}
\\ \\
\inference[\reappbad]
  {\pair{\stuck}{\su_2} = \force{v_1}{\tfun{\thole{}}{\thole{}}}}
  {\stepg{\eapp{v_1}{v_2}}{\su_1}{\tr}{\stuck}{\su_1\su_2}{\singlestep{\eapp{v_1}{v_2}}{\stuck}; \tr}}
\end{gather*}
\caption{A selection of the operational semantics from
  Figure~\ref{fig:operational}, extended to collect a full reduction
  graph.}
\label{fig:interactive}
\end{figure*}
%
The changes to the operational semantics of \S\ref{sec:semantics} are
mechanical, so we will not reproduce them in full; instead we describe
the procedure for extending a transition rule and provide a selection of
examples in Figure~\ref{fig:interactive}.

@@

\subsection{Traversing the Reduction Graph}
\label{sec:traversing-graph}
Given a reduction graph $G$ and an initial path
$p = \singlestep{e_1}{\singlestep{e_2}{e_n}}$, such that
\stepsg{e_1}{\emptysu}{\bullet}{e_n}{\su}{\tr}, we define the following
traversals in Figure~\ref{fig:traversing-graph}:
%
\begin{itemize}
\item \stepforwardsym takes a single step forward.
\item \stepbackwardsym takes a single step backward.
\item \jumpforwardsym takes a ``big'' step forward to the next function call.
\item \jumpbackwardsym takes a ``big'' step backward to the previous function call.
\item \stepintosym steps into a function call in a sub-term, isolating it from the context.
\item \stepoversym steps over a function call in a sub-term.
\end{itemize}
%
The initial path $p$ is required for the backward-steps as a node may
have multiple incoming $\leadsto$ edges, \eg
\singlestep{\eplus{1}{2}}{3} and \singlestep{\eplus{2}{1}}{3}.
%
The sub-term edges $\searrow$ allow us to decompose an expression into a
sub-expression and the surrounding context, thus enabling the \stepintosym
and \stepoversym traversals.

\begin{figure*}[t]
\centering
\[
\begin{array}{lcl}
\stepforward{e_i}  &\defeq& {e_{i+1}} \\ \\
\stepbackward{e_i} &\defeq& {e_{i-1}} \\ \\
\jumpforward{e_i} &\defeq& \left\{\begin{array}{ll}
                         e_{i+1}, & \text{if } e_{i+1} = \eapp{v_1}{v_2} \\
                         \jumpforward{e_{i+1}}, & \text{otherwise}
                         \end{array}\right\} \\ \\
\jumpbackward{e_i} &\defeq& \left\{\begin{array}{ll}
                         e_{i-1}, & \text{if } e_{i-1} = \eapp{v_1}{v_2} \\
                         \jumpbackward{e_{i-1}}, & \text{otherwise}
                         \end{array}\right\} \\ \\
\stepinto{e_i} &\defeq& \left\{\begin{array}{ll}
                         e\sub{x}{v_2}, & \text{if } e_i = C[\eapp{v_1}{v_2}] \text{ and } \singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}
                         \end{array}\right\} \\ \\
\stepover{e_i} &\defeq& \left\{\begin{array}{ll}
                         C[v], & \text{if } e_i = C[\eapp{v_1}{v_2}] \text{ and } \multistep{\eapp{v_1}{v_2}}{v}
                         \end{array}\right\}
\end{array}
\]
\caption{Rules for traversing the reduction graph. \stepintosym and \stepoversym
  require a traversal of the sub-term edges to decompose $e_i$ into the
  target expression \eapp{v_1}{v_2} and the context $C$.}
\label{fig:traversing-graph}
\end{figure*}
% \begin{itemize}
% \item extend operational semantics to collect reduction graph
% \item nodes are terms, edges indicate ``steps-to'' and ``sub-term'' relations
% \item visualize path through reduction graph
% \item expand edges to reveal more fine-grained steps (step/jump forward/backward)
% \item never lose context (unlike traditional debugger)
% \end{itemize}

>>>>

\subsection{Interactive Semantics}
\label{sec:inter-semant}
%
\begin{figure*}[t]
\relDescription{Trace Syntax}
$$
\begin{array}{rrcl}
  & \tr & ::= & \bullet \spmid \singlestep{e}{e}; \tr \spmid \subterm{e}{e}; \tr
\end{array}
$$
\\
\relDescription{\subtermssym}
\begin{gather*}
\begin{array}{lcl}
\subterms{\eapp{e_1}{e_2}}   & \defeq & \subterm{\eapp{e_1}{e_2}}{e_1}; \subterm{\eapp{e_1}{e_2}}{e_2} \\
\subterms{\eplus{e_1}{e_2}}   & \defeq & \subterm{\eplus{e_1}{e_2}}{e_1}; \subterm{\eplus{e_1}{e_2}}{e_2} \\
\subterms{\eif{e_1}{e_2}{e_3}}   & \defeq & \subterm{\eif{e_1}{e_2}{e_3}}{e_1}; \\
                                &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_2}; \\
                                &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_3} \\
\subterms{\elet{x}{e_1}{e_2}}   & \defeq & \subterm{\elet{x}{e_1}{e_2}}{e_1}; \\
                                &        & \subterm{\elet{x}{e_1}{e_2}}{e_2} \\
\subterms{\efun{x}{e}}       & \defeq & \subterm{\efun{x}{e}}{e} \\
\subterms{e}                 & \defeq & \bullet
\end{array}
\end{gather*}
\judgementHead{Evaluation}{\stepg{e}{\su}{\tr}{e}{\su}{\tr}}
\begin{gather*}
\inference[\recontext]
  {\stepg{e}{\su}{\tr}{e_1}{\su_1}{\tr_1}}
  {\stepg{C[e]}{\su}{\tr}{C[e_1]}{\su_1}{\singlestep{C[e]}{C[e_1]}; \subterms{C[e_1]}; \tr_1}}
\\ \\
\inference[\reappgood]
  {\pair{\efun{x}{e}}{\su_2} = \force{v_1}{\tfun{\thole{}}{\thole{}}}}
  {\stepg{\eapp{v_1}{v_2}}{\su_1}{\tr}
         {e\sub{x}{v_2}}{\su_1;\su_2}{\singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}; \subterms{e\sub{x}{v_2}}; \tr}}
\\ \\
\inference[\reappbad]
  {\pair{\stuck}{\su_2} = \force{v_1}{\tfun{\thole{}}{\thole{}}}}
  {\stepg{\eapp{v_1}{v_2}}{\su_1}{\tr}{\stuck}{\su_1;\su_2}{\singlestep{\eapp{v_1}{v_2}}{\stuck}; \tr}}
\end{gather*}
\caption{A selection of the operational semantics from
  Figure~\ref{fig:operational}, extended to collect a full reduction
  graph.}
\label{fig:interactive}
\end{figure*}
%
The changes to the operational semantics of \S\ref{sec:semantics} are
mechanical, so we will not reproduce them in full; instead we describe
the procedure for extending a transition rule and provide a selection of
examples in Figure~\ref{fig:interactive}.

@@

\subsection{Traversing the Reduction Graph}
\label{sec:traversing-graph}
Given a reduction graph $G$ and an initial path
$p = \singlestep{e_1}{\singlestep{e_2}{e_n}}$, such that
\stepsg{e_1}{\emptysu}{\bullet}{e_n}{\su}{\tr}, we define the following
traversals in Figure~\ref{fig:traversing-graph}:
%
\begin{itemize}
\item \stepforwardsym takes a single step forward.
\item \stepbackwardsym takes a single step backward.
\item \jumpforwardsym takes a ``big'' step forward to the next function call.
\item \jumpbackwardsym takes a ``big'' step backward to the previous function call.
\item \stepintosym steps into a function call in a sub-term, isolating it from the context.
\item \stepoversym steps over a function call in a sub-term.
\end{itemize}
%
The initial path $p$ is required for the backward-steps as a node may
have multiple incoming $\leadsto$ edges, \eg
\singlestep{\eplus{1}{2}}{3} and \singlestep{\eplus{2}{1}}{3}.
%
The sub-term edges $\searrow$ allow us to decompose an expression into a
sub-expression and the surrounding context, thus enabling the \stepintosym
and \stepoversym traversals.
%

\begin{figure*}[t]
\centering
\[
\begin{array}{lcl}
\stepforward{e_i}  &\defeq& {e_{i+1}} \\ \\
\stepbackward{e_i} &\defeq& {e_{i-1}} \\ \\
\jumpforward{e_i} &\defeq& \left\{\begin{array}{ll}
                         e_{i+1}, & \text{if } e_{i+1} = \eapp{v_1}{v_2} \\
                         \jumpforward{e_{i+1}}, & \text{otherwise}
                         \end{array}\right\} \\ \\
\jumpbackward{e_i} &\defeq& \left\{\begin{array}{ll}
                         e_{i-1}, & \text{if } e_{i-1} = \eapp{v_1}{v_2} \\
                         \jumpbackward{e_{i-1}}, & \text{otherwise}
                         \end{array}\right\} \\ \\
\stepinto{e_i} &\defeq& \left\{\begin{array}{ll}
                         e\sub{x}{v_2}, & \text{if } e_i = C[\eapp{v_1}{v_2}] \text{ and } \singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}
                         \end{array}\right\} \\ \\
\stepover{e_i} &\defeq& \left\{\begin{array}{ll}
                         C[v], & \text{if } e_i = C[\eapp{v_1}{v_2}] \text{ and } \multistep{\eapp{v_1}{v_2}}{v}
                         \end{array}\right\}
\end{array}
\]
\caption{Rules for traversing the reduction graph. \stepintosym and \stepoversym
  require a traversal of the sub-term edges to decompose $e_i$ into the
  target expression \eapp{v_1}{v_2} and the context $C$.
  \ES{need to mention $G$ and $p$ in these rules}
}
\label{fig:traversing-graph}
\end{figure*}

% \begin{itemize}
% \item extend operational semantics to collect reduction graph
% \item nodes are terms, edges indicate ``steps-to'' and ``sub-term'' relations
% \item visualize path through reduction graph
% \item expand edges to reveal more fine-grained steps (step/jump forward/backward)
% \item never lose context (unlike traditional debugger)
% \end{itemize}
